---
title: "Look at habitat structure as predictors for richness"
output: html_notebook
---

```{r}
library(tidyr)
library(dplyr)
library(readr)
library(MASS)
library(MuMIn)
library(ggplot2)
library(ggpubr)
library(cowplot)
options(na.action = "na.fail")
set.seed(123456)
```

# Calculate taxa richness per survey square
```{r}
species_per_grid = read_csv('./output/2_species_data_by_habitat_and_taxa_group.csv')
species_per_grid
```
```{r}
taxa_richness_per_grid = species_per_grid %>% filter(!is.na(title)) %>% dplyr::select(title, species, taxa_group) %>% distinct() %>% group_by(title, taxa_group) %>% dplyr::summarise(richness = n())
taxa_richness_per_grid
```
Check number of distinct survey grids
```{r}
taxa_richness_per_grid %>% dplyr::select(title) %>% distinct() %>% count()
```

```{r}
habitat_structure_predictors = read_csv('input/5/ground_cover_classification-1m.csv') %>% rename(gnd_cvr=herbaceous_coverage, shrb = shrub_coverage, yng_tr = young_tree_coverage, mat_tr = mature_tree_coverage, old_tr = old_tree_coverage)
habitat_structure_predictors
```


```{r}
create_data_for_taxa = function(taxa_name) {
  present_records = taxa_richness_per_grid %>% filter(taxa_group == taxa_name) %>% left_join(habitat_structure_predictors, by = 'title')
  
  absent_records = taxa_richness_per_grid %>% dplyr::select(title) %>% distinct() %>% filter(!(title %in% present_records$title)) %>% left_join(habitat_structure_predictors, by = 'title') %>% mutate(richness = 0, taxa_group = taxa_name)
  
  rbind(present_records, absent_records) %>% ungroup() %>% dplyr::select(-title) %>% dplyr::select(-taxa_group)
}
create_data_for_taxa('insect')
```

## Habitat structure for insects
```{r}
insect_data = create_data_for_taxa('insect')
insect_model = glm(richness ~ gnd_cvr + shrb + yng_tr + mat_tr + old_tr, gaussian, insect_data)
summary(insect_model)
```

```{r}
model_interactions <- lm(richness ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^2, insect_data)
summary(model_interactions)
```

```{r}
model_interactions <- lm(richness ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^3, insect_data)
summary(model_interactions)
```

```{r}
insect_model_selected = stepAIC(insect_model, scope = ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^5, direction = "both")
```

```{r}
summary(insect_model_selected)
```

```{r}
AIC(insect_model, insect_model_selected)
```

```{r}
insect_model_dredge = dredge(glm(richness ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^3, gaussian, insect_data))
insect_model_dredge_best = get.models(insect_model_dredge, 1)[[1]]
summary(insect_model_dredge_best)
```

```{r}
AIC(insect_model, insect_model_selected, insect_model_dredge_best)
```

## Habitat structure for birds

### From survey area
```{r}
bird_data = create_data_for_taxa('bird')
bird_model = glm(richness ~ gnd_cvr + shrb + yng_tr + mat_tr + old_tr, gaussian, bird_data)
summary(bird_model)
```

```{r}
bird_model_selected = stepAIC(bird_model, scope = ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^5, direction = "both")
```

```{r}
summary(bird_model_selected)
```

```{r}
AIC(bird_model, bird_model_selected)
```

```{r}
bird_model_dredge = dredge(glm(richness ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^3, gaussian, bird_data))
bird_model_dredge_best = get.models(bird_model_dredge, 1)[[1]]
summary(bird_model_dredge_best)
```

```{r}
AIC(bird_model, bird_model_selected, bird_model_dredge_best)
```

### From 50m buffer around survey area
```{r}
habitat_structure_predictors_from_buffer_in = read_csv('input/5/ground_cover_classification-1m-in_buffer.csv') %>% rename(gnd_cvr=herbaceous_coverage, shrb = shrub_coverage, yng_tr = young_tree_coverage, mat_tr = mature_tree_coverage, old_tr = old_tree_coverage)
habitat_structure_predictors_from_buffer_in
```
```{r}
habitat_structure_predictors_from_buffer_in %>% mutate(total_area = gnd_cvr + shrb + yng_tr + mat_tr + old_tr) %>% arrange(desc(total_area))
```

```{r}
scale = 400 / 12260
habitat_structure_predictors_from_buffer = habitat_structure_predictors_from_buffer_in %>% mutate(
  gnd_cvr = ceiling(gnd_cvr * scale),
  shrb = ceiling(shrb * scale),
  yng_tr = ceiling(yng_tr * scale),
  mat_tr = ceiling(mat_tr * scale),
  old_tr = ceiling(old_tr * scale))
habitat_structure_predictors_from_buffer
```

```{r}
create_data_from_buffer_for_birds = function() {
  taxa_name = 'bird'
  present_records = taxa_richness_per_grid %>% filter(taxa_group == taxa_name) %>% left_join(habitat_structure_predictors_from_buffer, by = 'title')
  
  absent_records = taxa_richness_per_grid %>% dplyr::select(title) %>% distinct() %>% filter(!(title %in% present_records$title)) %>% left_join(habitat_structure_predictors_from_buffer, by = 'title') %>% mutate(richness = 0, taxa_group = taxa_name)
  
  rbind(present_records, absent_records) %>% ungroup() %>% dplyr::select(-title) %>% dplyr::select(-taxa_group)
}
```

```{r}
bird_buf_data = create_data_from_buffer_for_birds()
bird_buf_model = glm(richness ~ gnd_cvr + shrb + yng_tr + mat_tr + old_tr, gaussian, bird_buf_data)
summary(bird_buf_model)
```

```{r}
bird_buf_model_selected = stepAIC(bird_buf_model, scope = ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^5, direction = "both")
```

```{r}
summary(bird_buf_model_selected)
```

```{r}
AIC(bird_buf_model, bird_buf_model_selected)
```

```{r}
bird_model_buf_dredge = dredge(glm(richness ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^3, gaussian, bird_buf_data))
bird_model_buf_dredge_best = get.models(bird_model_buf_dredge, 1)[[1]]
summary(bird_model_buf_dredge_best)
```

```{r}
AIC(bird_buf_model, bird_buf_model_selected, bird_model_buf_dredge_best)
```


## Habitat structure for plants
```{r}
plant_data = create_data_for_taxa('plant')
plant_model = glm(richness ~ gnd_cvr + shrb + yng_tr + mat_tr + old_tr, gaussian, plant_data)
summary(plant_model)
```

```{r}
plant_model_selected = stepAIC(plant_model, scope = ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^5, direction = "both")
```

```{r}
summary(plant_model_selected)
```

```{r}
AIC(plant_model, plant_model_selected)
```

```{r}
plant_model_dredge = dredge(glm(richness ~ (gnd_cvr + shrb + yng_tr + mat_tr + old_tr)^3, gaussian, plant_data))
plant_model_dredge_best = get.models(plant_model_dredge, 1)[[1]]
summary(plant_model_dredge_best)
```

```{r}
AIC(plant_model, plant_model_selected, plant_model_dredge_best)
```

# Plot dredge results
```{r}
.coefarr.avg = function(cfarr, weight, revised.var, full, alpha) {	
    weight <- weight / sum(weight)
    nCoef <- dim(cfarr)[3L]
    if(full) {
      nas <- is.na(cfarr[, 1L, ]) & is.na(cfarr[, 2L, ])
      cfarr[, 1L, ][nas] <- cfarr[, 2L, ][nas] <- 0
      #cfarr[, 1L:2L, ][is.na(cfarr[, 1L:2L, ])] <- 0
      if(!all(is.na(cfarr[, 3L, ])))
        cfarr[ ,3L, ][is.na(cfarr[ , 3L, ])] <- Inf
    }
    
    avgcoef <- array(dim = c(nCoef, 5L),
                     dimnames = list(dimnames(cfarr)[[3L]], c("Estimate",
                                                              "Std. Error", "Adjusted SE", "Lower CI", "Upper CI")))
    for(i in seq_len(nCoef))
      avgcoef[i, ] <- par.avg(cfarr[, 1L, i], cfarr[, 2L, i], weight,
                              df = cfarr[, 3L, i], alpha = alpha, revised.var = revised.var)
    
    avgcoef[is.nan(avgcoef)] <- NA
    return(avgcoef)
  }

.makecoefmat = function(cf) {
  no.ase <- all(is.na(cf[, 3L]))
  z <- abs(cf[, 1L] / cf[, if(no.ase) 2L else 3L])
  pval <- 2 * pnorm(z, lower.tail = FALSE)
  cbind(cf[, if(no.ase) 1L:2L else 1L:3L, drop = FALSE],
        `z value` = z, `Pr(>|z|)` = zapsmall(pval))
}

model_average <- function(dredge_result) {
  summary(model.avg(dredge_result))
}

dredge_summary <- function(model_sum) {
  # just return the estimate and p value
  weight <- model_sum$msTable[, 5L]
  
  coefmat.full <- as.data.frame(.makecoefmat(.coefarr.avg(model_sum$coefArray, weight,
                                                          attr(model_sum, "revised.var"), TRUE, 0.05)))
  coefmat.full <- coefmat.full[-c(1), c(1, 2, 5)]
  names(coefmat.full) <- c("estimate", "error", "p")
  coefmat.full <- tibble::rownames_to_column(coefmat.full, "explanatory")
  
  coefmat.full
}

plot_vsurp_result = function(taxa_name, result_table, limit) {
  neg_limit = -1 * limit
  
  ggplot(result_table, aes(y=explanatory, x=estimate)) + 
    geom_line() +
    geom_point()+
    geom_errorbar(aes(xmin=estimate-error, xmax=estimate+error), width=.2,
                   position=position_dodge(0.05)) +
    geom_vline(xintercept=0, linetype="dotted") +
    xlab('Increase in proportion of group presence\n± Standard Error') + ylab('Predictor') +
    labs(title = taxa_name) +
    scale_x_continuous(limits = c(neg_limit, limit)) +
    theme_bw() +
    theme(legend.justification = "top")
}
```

## Insects
```{r}
insect_model_avg = model_average(insect_model_dredge)
insect_sum = dredge_summary(insect_model_avg)
```

```{r}
plot_vsurp_result('insect', insect_sum, 0.025)
```

## Birds (within buffer)
```{r}
bird_buf_model_avg = model_average(bird_model_buf_dredge)
bird_buf_sum = dredge_summary(bird_buf_model_avg)
```

```{r}
plot_vsurp_result('bird', bird_buf_sum, 0.05)
```


## Plants
```{r}
plant_model_avg = model_average(plant_model_dredge)
plant_sum = dredge_summary(plant_model_avg)
```

```{r}
plot_vsurp_result('plant', plant_sum, 0.025)
```

# Explore interactions

## Insects
```{r}
summary(insect_model_dredge_best)
```
Call:
glm(formula = richness ~ gnd_cvr + mat_tr + shrb + gnd_cvr:mat_tr + 
    gnd_cvr:shrb + mat_tr:shrb + gnd_cvr:mat_tr:shrb + 1, family = gaussian, 
    data = insect_data)

Coefficients:
                   Estimate Std. Error t value Pr(>|t|)    
(Intercept)       4.423e-01  1.113e-01   3.973  8.4e-05 ***
gnd_cvr              2.889e-04  3.908e-04   0.739 0.460146    
mat_tr            3.851e-03  1.310e-03   2.939 0.003479 ** 
shrb              7.502e-03  5.419e-03   1.384 0.167059    
gnd_cvr:mat_tr      -3.530e-05  1.040e-05  -3.393 0.000759 ***
gnd_cvr:shrb        -1.662e-05  2.339e-05  -0.710 0.477830    
mat_tr:shrb      -9.927e-05  4.992e-05  -1.989 0.047428 *  
gnd_cvr:mat_tr:shrb  1.015e-06  3.732e-07   2.720 0.006820 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

(Dispersion parameter for gaussian family taken to be 0.6420709)

    Null deviance: 274.16  on 409  degrees of freedom
Residual deviance: 258.11  on 402  degrees of freedom
AIC: 991.8

Number of Fisher Scoring iterations: 2

```{r}
habitat_structure_predictors_insect_pred = habitat_structure_predictors
habitat_structure_predictors_insect_pred$insect_richness = predict(insect_model_dredge_best, habitat_structure_predictors_insect_pred, type="response")
```


```{r}
min(habitat_structure_predictors_insect_pred$mat_tr)
max(habitat_structure_predictors_insect_pred$mat_tr)
```

```{r}
mat_tr_bin_edges <- c(0, 100, 200, 300, 400)
habitat_structure_predictors_insect_pred$mat_tr_bin = cut(habitat_structure_predictors_insect_pred$mat_tr, breaks = mat_tr_bin_edges, include.lowest = TRUE)
```


```{r}
ggplot(habitat_structure_predictors_insect_pred, aes(x = gnd_cvr, y = insect_richness, color = factor(mat_tr_bin))) +
  geom_smooth(size = 1) +
  labs(title = "Interaction effect of ground cover and shrubs on insect_richness",
       x = "gnd_cvr",
       y = "Predicted insect richness",
       color = "mat_tr Level") +
  theme_minimal()
```


```{r}
ggplot(habitat_structure_predictors_insect_pred, aes(x = shrb, y = insect_richness, color = factor(mat_tr_bin))) +
  geom_smooth(size = 1) +
  labs(title = "Interaction effect of ground cover and shrubs on insect_richness",
       x = "shrb",
       y = "Predicted insect richness",
       color = "mat_tr Level") +
  theme_minimal()
```
```{r}
label_percent_of_400 = function(x) paste0(x*0.25, "%")
```

gnd_cvr:mat_tr      -3.530e-05  1.040e-05  -3.393 0.000759 ***
gnd_cvr:shrb        -1.662e-05  2.339e-05  -0.710 0.477830    
mat_tr:shrb      -9.927e-05  4.992e-05  -1.989 0.047428 * 
```{r}
habitat_structure_predictors_insect_pred2 <- expand.grid(
  mat_tr = seq(min(habitat_structure_predictors$mat_tr), max(habitat_structure_predictors$mat_tr), length.out = 50),
  shrb = seq(min(habitat_structure_predictors$shrb), max(habitat_structure_predictors$shrb), length.out = 50),
  gnd_cvr = seq(min(habitat_structure_predictors$gnd_cvr), max(habitat_structure_predictors$gnd_cvr), length.out = 50)
)

habitat_structure_predictors_insect_pred2 = habitat_structure_predictors_insect_pred2 %>%
  filter(mat_tr + shrb + gnd_cvr <= 400)

habitat_structure_predictors_insect_pred2$insect_richness = predict(plant_model_dredge_best, habitat_structure_predictors_insect_pred2, type="response")

p1 = ggplot(habitat_structure_predictors_insect_pred2, aes(x = mat_tr, y = gnd_cvr, fill = insect_richness)) +
  geom_tile() +
  scale_fill_viridis_c() +  # Better color scaling
  labs(
       x = "Mature tree cover", 
       y = "Ground cover", 
       fill = "Predicted insect richness") +
  scale_y_continuous(label = label_percent_of_400) +
  scale_x_continuous(label = label_percent_of_400) +
  theme_minimal()

p2 = ggplot(habitat_structure_predictors_insect_pred2, aes(x = shrb, y = gnd_cvr, fill = insect_richness)) +
  geom_tile() +
  scale_fill_viridis_c() +  # Better color scaling
  labs(
       x = "Shrub cover", 
       y = "Ground cover", 
       fill = "Predicted insect richness") +
  scale_y_continuous(label = label_percent_of_400) +
  scale_x_continuous(label = label_percent_of_400) +
  theme_minimal()

p3 = ggplot(habitat_structure_predictors_insect_pred2, aes(x = mat_tr, y = shrb, fill = insect_richness)) +
  geom_tile() +
  scale_fill_viridis_c() +  # Better color scaling
  labs(
       x = "Mature tree cover", 
       y = "Shrub cover", 
       fill = "Predicted insect richness") +
  scale_y_continuous(label = label_percent_of_400) +
  scale_x_continuous(label = label_percent_of_400) +
  theme_minimal()

legend <- ggpubr::get_legend(
  # create some space to the left of the legend
  p1
)

title <- ggdraw() + 
  draw_label(
    "Interaction effects of ground cover, shrbs and mature trees on insect richness",
    fontface = 'bold',
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )

pg = plot_grid(
  nrow = 2, ncol = 2,
  p1 + theme(legend.position="none"),
  p2 + theme(legend.position="none"),
  p3 + theme(legend.position="none"),
  legend
)
plot_grid(nrow = 2, ncol = 1, title, pg, rel_heights = c(0.1, 1))
ggsave('output/5/insect_predictor_interactions.jpg')
```


## Plants

```{r}
summary(plant_model_dredge_best)
```
glm(formula = richness ~ gnd_cvr + mat_tr + shrb + gnd_cvr:shrb + 1, 
    family = gaussian, data = plant_data)

Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)  2.320e+00  1.278e-01  18.154  < 2e-16 ***
gnd_cvr        -1.159e-04  4.499e-04  -0.258    0.797    
mat_tr       4.145e-03  8.603e-04   4.818 2.05e-06 ***
shrb         2.319e-03  3.705e-03   0.626    0.532    
gnd_cvr:shrb    4.803e-05  1.926e-05   2.495    0.013 *  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

(Dispersion parameter for gaussian family taken to be 0.9659682)

    Null deviance: 458.39  on 409  degrees of freedom
Residual deviance: 391.22  on 405  degrees of freedom
AIC: 1156.3

Number of Fisher Scoring iterations: 2

```{r}
habitat_structure_predictors_plant_pred = habitat_structure_predictors
habitat_structure_predictors_plant_pred$plant_richness = predict(plant_model_dredge_best, habitat_structure_predictors_plant_pred, type="response")
```


```{r}
min(habitat_structure_predictors_plant_pred$shrb)
max(habitat_structure_predictors_plant_pred$shrb)
```

```{r}
shrb_bin_edges <- c(0, 50, 100, 150, 200, 257)
habitat_structure_predictors_plant_pred$shrb_bin = cut(habitat_structure_predictors_plant_pred$shrb, breaks = shrb_bin_edges, include.lowest = TRUE)
```


```{r}
ggplot(habitat_structure_predictors_plant_pred, aes(x = gnd_cvr, y = plant_richness, color = factor(shrb_bin))) +
  geom_smooth(size = 1) +
  labs(title = "Interaction effect of ground cover and shrubs on plant richness",
       x = "gnd_cvr",
       y = "Predicted plant richness",
       color = "Shrub cover") +
  theme_minimal()
```

```{r}
habitat_structure_predictors_plant_pred2 <- expand.grid(
  mat_tr = seq(min(habitat_structure_predictors$mat_tr), max(habitat_structure_predictors$mat_tr), length.out = 50),
  shrb = seq(min(habitat_structure_predictors$shrb), max(habitat_structure_predictors$shrb), length.out = 50),
  gnd_cvr = seq(min(habitat_structure_predictors$gnd_cvr), max(habitat_structure_predictors$gnd_cvr), length.out = 50)
)

habitat_structure_predictors_plant_pred2 = habitat_structure_predictors_plant_pred2 %>%
  filter(mat_tr + shrb + gnd_cvr <= 400)

habitat_structure_predictors_plant_pred2$plant_richness = predict(plant_model_dredge_best, habitat_structure_predictors_plant_pred2, type="response")

ggplot(habitat_structure_predictors_plant_pred2, aes(x = gnd_cvr, y = shrb, fill = plant_richness)) +
  geom_tile() +
  scale_fill_viridis_c() +  # Better color scaling
  labs(title = "Interaction effect of ground cover and shrubs on plant richness",
       x = "gnd_cvr", 
       y = "shrb", 
       fill = "Predicted plant richness") +
  scale_y_continuous(label = label_percent_of_400) +
  scale_x_continuous(label = label_percent_of_400) +
  xlab('Ground cover') + ylab("Shrub cover") +
  theme_minimal()
ggsave('output/5/plant_predictor_interactions.jpg')
```

## Birds - but including 50m buffer around survey area

```{r}
summary(bird_model_buf_dredge_best)
```

