---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(dplyr)
library(dbplyr)
library(readxl)
```

```{r}
data = read_csv('./input/included_species.csv')
data
```

```{r}
avonet = read_csv('./input/AVONET2_eBird.csv')
avonet
```

```{r}
bird_avonet_habitat = data %>% filter(group == 'bird') %>% left_join(avonet, by = c('scientific name' = 'Species2')) %>% select('common name', 'scientific name', Habitat, group) %>% rename(common_name='common name', scientific_name='scientific name', habitat = Habitat)

bird_avonet_habitat$habitat[bird_avonet_habitat$habitat == 'Forest'] = 'Woodland'
bird_avonet_habitat$habitat[bird_avonet_habitat$habitat == 'Human Modified'] = 'Built-Up'
bird_avonet_habitat$habitat[bird_avonet_habitat$habitat == 'Riverine'] = 'Wetlands'

bird_avonet_habitat
```
```{r}
birdlife_data = read_excel('./input/birdlife_datazone_habitats.xlsx')
birdlife_data
```

```{r}
bird_birdlife_habitat = data %>% filter(group == 'bird') %>% left_join(birdlife_data) %>% select('common name', 'scientific name', habitat, group) %>% rename(common_name='common name', scientific_name='scientific name')
bird_birdlife_habitat
```
```{r}
bird_habitat = distinct(rbind(bird_avonet_habitat, bird_birdlife_habitat))
bird_habitat
```

```{r}
plant_data = read_excel('./input/PLANTATT_19_Nov_08.xls')
plant_data %>% select('Taxon name', 'Fam', 'NS', 'Hght', 'GB', 'Br Habitats') %>% right_join(data %>% filter(group == 'plant'), by = c('Taxon name' = 'scientific name'))
```

Broad Habitat codes
1   Broadleaved, mixed and yew woodland
2   Coniferous woodland
3   Boundary and linear features (eg hedges, roadsides, walls)
4   Arable and horticultural (includes orchards, excludes domestic gardens)
5   Improved grassland
6   Neutral grassland (includes coarse Arrhenatherum grassland)
7   Calcareous grassland (includes lowland and montane types)
8   Acid grassland (includes non-calcareous sandy grassland)
9   Bracken
10   Dwarf shrub heath (cover of dwarf shrubs at least 25%)
11   Fen, marsh and swamp (not wooded)
12   Bog
13   Standing water and canals
14   Rivers and streams
15   Montane habitats (acid grassland and heath with montane species)
16   Inland rock (quarries, cliffs, screes)
17   Built-up areas and gardens
18   Supralittoral rock (does not include maritime grassland)
19   Supralittoral sediment (strandlines, shingle, coastal dunes)
21   Littoral sediment (includes saltmarsh and saltmarsh pools)
23   Inshore sublittoral sediment (only Zostera marina)
```{r}
plant_habitat_dict = data.frame(
  habitat_code = 1:23,
  habitat = c('Woodland', 'Woodland', 'Boundary', 'Arable', 'Grassland', 'Grassland', 'Grassland', 'Grassland', 'Bracken', 'Shrub', 'Wetlands', 'Wetlands', 'Wetlands', 'Wetlands', 'Mountain', 'Rock', 'Built-Up', 'Rock', 'Sediment', NA, 'Saltmarsh', NA, 'Inshore')
  )
plant_habitat_dict
```

```{r}
plant_habitat = plant_data %>% select('Taxon name', 'Br Habitats') %>% rename(habitat_code = 'Br Habitats') %>% right_join(data %>% filter(group == 'plant'), by = c('Taxon name' = 'scientific name')) %>% separate_rows(habitat_code, convert = TRUE, sep = ",") %>% left_join(plant_habitat_dict, by = c('habitat_code' = 'habitat_code')) %>% select('common name', 'Taxon name', 'habitat', group) %>% rename(common_name='common name', scientific_name='Taxon name')
plant_habitat 
```

```{r}
insect_data = read_excel('./input/wildlifetrust_habitats.xlsx')
insect_data
```

```{r}
insect_habitat = insect_data %>% right_join(data %>% filter(group == 'insect')) %>% select('common name', 'scientific name', 'habitat', group) %>% rename(common_name='common name', scientific_name='scientific name')
insect_habitat
```
```{r}
rbind(bird_habitat, insect_habitat, plant_habitat) %>% select(habitat) %>% filter(!is.na(habitat)) %>% distinct()
```

```{r}
species_to_habitat = rbind(bird_habitat, insect_habitat, plant_habitat) %>% filter(habitat %in% c('Woodland', 'Wetlands', 'Built-Up', 'Arable', 'Rock', 'Shrub', 'Grassland'))
species_to_habitat
```

```{r}
table(species_to_habitat$group, species_to_habitat$habitat)
```

```{r}
write_csv(species_to_habitat, './output/species_by_habitat.csv')
```
